### 2022-10-25 src/tex/ms.tex
add in inference references

% Define document class
\documentclass[twocolumn]{aastex631}

\newcommand{\flatiron}{\affiliation{Center for Computational Astrophysics,
Flatiron Institute, New York, NY 10010, USA}}
\newcommand{\stonybrook}{\affiliation{Department of Physics and Astronomy, Stony
Brook University, Stony Brook NY 11794, USA}}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{xcolor}
\usepackage{showyourwork}
\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
\newcommand{\wf}[1]{\textcolor{cyan}{WF: #1}}
\newcommand{\kb}[1]{\textcolor{pink}{#1}}
% \setlength{\floatsep}{0cm} \setlength{\textfloatsep}{0cm}
% \setlength{\belowcaptionskip}{-5pt} \setlength{\abovecaptionskip}{-5pt} Begin!
\begin{document}
\title{Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors}

\pacs{}

\author{Kaze W. K. Wong}
\email{kwong@flatironinstitute.org}
\flatiron

\author{Katelyn Breivik}
\email{kbreivik@flatironinstitute.org}
\flatiron

\author{Will M. Farr}
\email{will.farr@stonybrook.edu}
\flatiron
\stonybrook

\author{Rodrigo Luger}
\email{rluger@flatironinstitute.org}
\flatiron

\date{\today}

\begin{abstract}
One promising way to extract information about stellar astrophysics from
gravitational wave catalogs is to compare the catalog to the outputs of stellar
population synthesis modeling with varying physical assumptions.  The parameter
space of physical assumptions in population synthesis is high-dimensional and
the choice of parameters that best represents the evolution of a binary system
may depend in an as-yet-to-be-determined way on the system's properties.  Here
we propose a pipeline to simultaneously infer zero-age main sequence properties
and population synthesis parameter settings controlling modeled binary evolution
from individual gravitational wave observations of merging compact binaries. Our
pipeline can efficiently explore the high-dimensional space of population
synthesis settings and progenitor system properties for each system in a catalog
of gravitational wave observations.  We apply our pipeline to observations in
the third LIGO Virgo Kagra Gravitational Wave Transient Catalog. We showcase the
effectiveness of this pipeline with a detailed study of the progenitor
properties and population synthesis settings that produce mergers like the
observed GW150914.  Our pipeline permits a measurement of the variation of
population synthesis parameter settings with binary properties, if any; we
present inferences for the recent GWTC-3 transient catalog that suggest that the
stable mass transfer efficiency parameter may vary with primary black hole mass.  
\end{abstract}

%\maketitle

\section{Introduction}
As the detection rate of gravitation waves (GWs) from merging
double-compact-object (DCO) binaries increases with the sensitivity of the
ground-based GW detector network
\citep{Aso2013,LIGOScientificCollaboration2015,Acernese2015,Abbott2018,Buikema2020,Tse2019,Acernese2019,Akutsu2021},
we are beginning to constrain the astrophysical processes which shape the
evolution of GW progenitor populations. One of the most common ways to study
progenitor populations of GW mergers is through population synthesis simulations
of stellar populations from a host of formation environments. In the case of
isolated binary star evolution, merging double compact objects can form from
massive binary stars through the standard stable mass transfer or common
envelope channels
\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022}
as well as through chemically homogenous evolution \citep{Mandel2016,
Marchant2016, deMink2016} or with population III stars
\citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017,
Tanikawa2021, Tanikawa2022}. Merging DCO binaries can also originate from a wide
variety of dynamically active environments including triple (or higher multiple)
systems \citep[e.g.][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021},
as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017,
DiCarlo2020, Chattopadhyay2022}, globular clusters
\citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014,
Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019}, nuclear star clusters
\citep[][]{Miller2009, Antonini2016}, or a mix of all three
\citep[e.g.][]{Mapelli2022}. Finally, more exotic channels like active galactic
nuclei which combine both gravitational and gas interactions
\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial
black holes \citep[e.g.][]{Bird2016, AliHaimoud2017} can produce merging DCOs
observable with a ground-based detector network. For a discussion of the
relative rates of each formation environment and channels within, see
\citet{Mandel2022} and references therein.

Traditionally, population synthesis studies simulate DCO populations with a
Monte-Carlo approach. Using theoretically motivated distributions of initial
parameters like age, metallicity, component mass, orbital separation, and
eccentricity, initial populations are evolved with fixed astrophysical
assumptions, or population synthesis hyperparameter settings, to produce a
synthetic catalog of DCO mergers which can be compared to parameterized models
derived from observations. \textbf{This method has been applied extensively to
inference on GW data to inform phenomenological models
\citep[e.g.][]{Fishbach2017, Wysocki2019, Callister2021, Delfavero2021,
Farah2022, Delfavero2022}, population synthesis for a single formation
environment \citep[e.g.][]{Belczynski2016, Stevenson2017b, Taylor2018,
Barrett2018, Mastrogiovanni2022} or contributions from multiple formation
environments \citep[e.g][]{Zevin2017, Stevenson2017, Bouffanais2019, Zevin2021,
Wong2021, Bouffanais2021, Mapelli2022}}. These approaches rely on several
assumptions, both in the simulations and model parameterizations fitted to the
GW data, which are unlikely to be fully correct. For example, there is no reason
to suppose that the entire population of DCOs should evolve with the same fixed
hyperparameters; there is no a priori reason the fitted parameterized model
should capture the relevant features of the observed DCO distributions to
correspond to the binary physics of interest; etc. A better approach would be to
compare progenitor parameters and hyperparameter settings directly to DCO
observations, treating the population synthesis model as a mapping from
progenitor to merger parameters; merger parameters can then be mapped into
observations using a gravitational waveform family, and parameter inference
proceeds in the usual way, propagating information back up the mapping chain
\citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
hyperparameter settings, was advanced in \citet{Andrews2018,Andrews2021}, as we
discuss in more detail below.

%One of the most common skepticisms of population synthesis studies is that they
%employ simulation sets which often cover a small subdomain of the entire
%parameter space of uncertain physics, or only vary one uncertain process per
%simulation (e.g. common envelope efficiency) while holding all others fixed to
%some ``default" values. On top of this, these values are often not calibrated
%to data.
Attempting to approach such an improved inference procedure for evolutionary
hyperparameters using traditional en mass Monte Carlo simulations of entire
populations of DCO mergers is computationally infeasible. Even with recent
developments in using emulators to speed up the simulation process
\citep[e.g.][]{Wong2021}, training the emulators still requires a significant
amount of computation to cover a wide range of uncertainty in the hyperparemeter
space. As an example, to emulate a model with $10$ parameters, the simplest way
to construct a training set of simulations is to run simulations on a grid which
varies combinations of uncertain physics. For $10$ uncertain physical processes,
if we consider $2$ variations for each physical process this would still require
$1024$ simulations. Because population synthesis simulations require hundreds to
thousands of cpu hours, training an emulator which spans a high dimensional
uncertainty space remains an impossibility at present.  Thus inference of
hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
systems via population synthesis recipes to match a particular observation one-
or few-at-a-time while simultaneously adjusting the hyperparameters. (Eventually
it may be even be possible to use more physically-motivated modeling of binary
physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)

\citet{Andrews2021} (hereafter A21) used \texttt{DartBoard} \citep{Andrews2018}
to determine the ZAMS parameters which produce GW150914-like BBH merger based on
posterior samples for GW150914 and a fixed set of hyperparameters which define
assumptions for how isolated binary-star interactions proceed using
\texttt{COSMIC}, a binary population synthesis code \citep{Breivik2020}. The
approach is very similar to the one advocated here \emph{except} that
\citet{Andrews2021} treated the evolutionary hyperparameters for the binary
physics as fixed within each analysis.  The key extension in this work, which
will enable population modeling of \emph{both} progenitor parameters and
hyperparemeter settings, including any possible \emph{dependence} of
hyperparameter settings on progenitor parameters, is to allow these
hyperparameters to vary at the same time as intrinsic properties of the
progenitor system to produce a joint inference over the intrinsic properties of
the progenitor and the necessary hyperparameter settings to produce the observed
merger properties.

In full: we propose a method to ``backward" model each GW event to its
progenitor state while allowing the hyperparameters to vary across their full
range of physical uncertainty following a two-stage process. First, we solve a
root finding problem to obtain guesses that are likely to produce the desired
system properties in the joint progenitor--hyperparameter space. We then sample
the posterior in the joint space that is induced by the DCO merger observations
using a Markov chain Monte Carlo (MCMC) algorithm that is initialized by the
roots found in the previous stage.  We demonstrate our algorithm in a realistic
setting by producing a posterior over progenitor properties and \texttt{COSMIC}
hyperparameters implied by the observation of GW150914.

We find that the progenitor properties and even the ability to produce a
GW150914-like merger event is strongly correlated with hyperparameter settings;
different assumptions about the black hole masses at merger in GW150914 imply
wildly different formation channels and ZAMS progenitor masses for this event
and some nearby combinations of ZAMS masses are unable, with any reasonable
hyperparameter settings, to produce a DCO merger like GW150914.  We further
exhibit preliminary results of an ongoing analysis over the entire GWTC-3
catalog \citep{GWTC-3} that suggest that the accretion efficiency in stable mass
transfer may depend non-trivially on the primary black hole mass in merging
systems; our methodology allows for a systematic study of such dependencies,
which we will explore more fully in future work.

The rest of the paper is structured as follows: We detail our method in Sec.
~\ref{sec:method}. In Sec. ~\ref{sec:result}, we show the results of applying
our method to a number of real events. We discuss the implications of this work
and future direction in Sec. ~\ref{sec:discussion}.

\section{Method}
\label{sec:method}

We define the progenitor's zero age main sequence (ZAMS) properties like mass,
orbital period, and eccentricity as progenitor parameters $\bm{\theta'}$, the
parameters that control different uncertain physical prescriptions such as wind
strength and common envelope efficiency as hyperparameters $\bm{\lambda}$, and
random variables that affect certain stochastic process such as whether CO natal
kick unbinds a binary as $\bm{X}$. To avoid clutter in the following derivation,
we denote all the parameters related to mapping a particular progenitor system
into a GW event collectively as evolutionary parameters $\bm{\Theta}$, which
includes $\bm{\theta'}$, $\bm{\lambda}$, and $\bm{X}$.

Once all the parameters including a particular draw of all the random variables
represented by $\bm{X}$ are known, the population synthesis code is a
deterministic function that transforms the properties of the progenitors into
the GW parameters $\bm{\theta}$:
\begin{equation}
    \bm{\theta} = \bm{f}\left( \bm{\Theta} \right).
\end{equation}
The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one due
to degeneracies in the different physical processes and initial parameters of
the GW event progenitors. In order to draw inferences about $\bm{\Theta}$ from
GW data, we must be able to evaluate the likelihood of that data at fixed
progenitor parameters, namely
\begin{equation}
    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right).
\end{equation}
The equality holds because the likelihood depends only on the gravitational
waveform generated by parameters $\bm{\theta}$ onto which the progenitor
parameters map.  In principle this likelihood could be computed at arbitrary
values of $\bm{\Theta}$ using the same machinery that is used to estimate source
parameters in GW catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3}
to develop samples of progenitor parameters, hyperparameters, and random
variables in $\bm{\Theta}$ from a posterior density.

Since the GW likelihood function ultimately depends only on the GW parameters
$\theta$, and GWTC-3 already has samples over these parameters from a posterior
density
\begin{equation}
    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) \pi_\mathrm{GW} \left( \bm{\theta} \right),
\end{equation}
where $\pi_\mathrm{GW}$ is the prior density used for sampling, we can save the
computational cost associated with evaluating the GW likelihood by rewriting the
posterior density of $\bm{\Theta}$ in terms of the posterior density of
$\bm{\theta}$, namely

\begin{align}
    p(\bm{\Theta} | d) = \frac{p(\bm{\theta}(\bm{\Theta})| d) \pi(\bm{\Theta})}{\pi_\mathrm{GW}(\bm{\theta}(\bm{\Theta}))}.
\end{align}
We apply a kernel density estimator with a Gaussian kernel to the posterior
samples released in GWTC-3 to estimate $p(\bm{\theta}(\bm{\Theta})|d)$.

Including hyperparameters and random variables %in the sampling
drastically increases the dimensionality of the problem, which can make the
sampling process much more computationally expensive to converge. To speed up
the convergence, we first solve a root-finding problem to find points
$\bm{\Theta}$ that will generate GW parameters $\bm{\theta}$ close to the bulk
of the posterior density of a GW event.  We then use those points to initialize
a set of chains in the MCMC process.

%The root-finding procedure is described as follows:
For each posterior sample point in the GW-observable space, we can find the
corresponding evolutionary parameters by solving an optimization problem that
minimizes the mean square difference between the GW event parameters and
evolutionary parameters as:

\begin{align}
\mathcal{L}(\bm{\Theta},\bm{\theta}) = ||f(\bm{\Theta})-\bm{\theta}||^2.
\label{eq:loss}
\end{align}

In principle, we should only accept solutions that exactly reproduce the LVK
posterior samples in the GW-observable space. However, it is not feasible to
achieve such a condition in practice, therefore we relax the condition in
eq.~\ref{eq:loss} to a small acceptance threshold. For this study, we picked a
threshold of $10^{-2}$. To make sure we find a reasonably complete set of
progenitor parameters that corresponds to the posterior sample point, we use
1000 different initial guesses in solving the optimization problem. As long as
the solution fulfills the acceptance criteria, the root is as valid as all other
roots that fulfill the same acceptance criteria, \emph{regardless to its initial
guess}. This means, we have the freedom to choose the set of initial guesses
however we think might benefit the optimization.

Most of the points in the evolutionary parameter space do not produce DCO
mergers in a Hubble time. For systems that do not merge, we set the binary
masses to be $0$ such that the gradient for the same systems is also $0$ and
leads the root finder to get stuck on its first step.{\footnote{On a plateau of
constant loss, the root finder performs no better than a random guess. Because
of this, a root finder in a high dimensional space will require a significant
amount of computation to get close to a solution.}} Therefore, it is beneficial
to choose initial guesses in region of the evolutionary parameter space that are
likely to produce DCO mergers. To construct a list of initial guesses, we evolve
a set of ZAMS parameters uniformly sampled from the initial binary parameter
space, then keep the systems that merge within a Hubble time.

Retaining explicit control over random variables allows us to marginalize over
their contribution, and thus allows us to focus on progenitor parameters and
hyperparameters. However, in practice most population synthesis codes (including
\texttt{COSMIC} which we use in this study) define their random variables
implicitly such that the random variables are drawn within the program instead
of being passed as arguments. This means that the process of evolving the binary
is not fully deterministic. Thus, even if the root-finding algorithm performs
perfectly, forward modelling a set of roots does not guarantee that the
simulated population reproduces exactly the set of posterior samples due to
randomness in the evolution of each binary. To assure that the recovered
progenitor and hyperparameters robustly correspond to the posterior in the
GW-observable space, we push forward (or ``reproject'') the recovered
evolutionary parameters to the GW-observable space to check whether the
reprojected posterior agrees with the posterior given by the LVK collaboration.

% Validity paragraph \wf{What if we made this a subsection of ``methods'', and
%then just one sentence here about the values of the KL divergence for
%GW150914?} To check the validity of the recovered posterior over evolutionary
%parameters, it is essential to check whether it is consistent with the original
%posterior in the observable space. We reproject the recovered posterior to the
%observable space, and compare the KL divergence between the original posterior
%and the reprojected posterior.

We use KL divergence to measure the agreement between the two posterior
distributions as:
\begin{align}
D_{KL}(P||Q) = \int P(\bm{x}) \log(P(\bm{x})/Q(\bm{x})) dx.
\label{eq:KLdivergence}
\end{align}

\noindent A small KL divergence means the reprojected \texttt{COSMIC} posterior
is similar to the original posterior in the observable space, and is thus a
viable channel for that specific event. Otherwise, it either means
\texttt{COSMIC} can only explain part of the posterior or cannot explain the
event at all. This is completely expected behavior since \texttt{COSMIC}, and
isolated binary evolution more generally, carries its own assumptions and is
expected to fail in reproducing a subset of the event in GWTC-3. One example of
this is GW events with at least one component with mass above the pair
instability supernova mass limit \citep{Woosley2017, Farmer2020}. The
reprojection could also be subject to stochasticity in the evolution of each
binary. To account for this, we reproject the posterior multiple times with
different random seeds and check whether the KL divergence varies significantly.
A varying KL divergence means a particular GW event is subject to randomness in
the evolution of the progenitor, and extra caution should be used when
interpreting the result.

\section{Results}
\label{sec:result}

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l }
    \hline
    \hline
    Parameters &  Description & Optimization range\\
    \hline
    \hline
    Observables &\ &\  \\
    \hline
    \hline
    $m_{\rm 1, GW}$ & Primary mass of the GW event & NA \\
    $m_{\rm 2, GW}$ & Secondary mass of the GW event  & NA\\
    \hline
    \hline
    Progenitor parameters &\ &\  \\
    \hline
    \hline
    $m_{\rm 1, ZAMS}$ & Primary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $m_{\rm 2, ZAMS}$ & Secondary mass at ZAMS & $[10,150]\ M_{\odot}$\\
    $t_{\rm orb}$ & Orbital period at ZAMS & $[5,5000]$ days\\
    $e$ & Eccentricity at ZAMS & $[0,1]$\\
    $Z$ & Metallicity at ZAMS & $[10^{-4},2\times10^{-2}]$\\
    $z$ & Redshift at ZAMS & NA\\
    \hline
    \hline
    hyperparameters &\ &\ \\
    \hline
    \hline

    $\alpha$ & Common envelope efficiency & $[0.1,10]$\\
    $f_{\rm acc}$ & Fraction of accretion during stable mass transfer &
    $[0,1]$\\
    $q_{\rm crit,3}$ & Critical mass ratio on the Hertzsprung Gap & $[0.1,10]$\\
    $\sigma$ & Root mean square of Maxwellian natal kick distribution&
    $[10,300]\ {\rm km/s}$\\

    \hline
    \hline
    \end{tabular}
    \caption{A list of parameters used in this study.}
    \label{tab:parameters}
    \end{center}
\end{table*}

We apply the method described in section \ref{sec:method} to the first GW event
GW150914. The parameters involved in the analysis and the range we allow for the
inference are tabulated in table \ref{tab:parameters}. The spin of BHs formed in
isolated binaries depends strongly on the angular momentum transport in massive
stars which is still highly uncertain (see however \citet{Fuller2019,
Bavera2020}), therefore we only consider the two component masses in the
observable space in both the root-finding stage and the MCMC stage. For
progenitor parameters, we characterize each guess with five parameters: the two
component masses $M_{\rm 1,ZAMS}$ and $M_{\rm 2,ZAMS}$, the orbital period
$t_{\rm orb}$, the eccentricity $e$, and the metallicity $Z$ at ZAMS. Note that
the progenitor formation redshift is not obtained through the root-finding
process because the formation redshift does not intrinsically affect the
evolution of the binary. Instead, we compute the time it takes for the binary to
evolve from formation to the merger and add this time to the lookback time when
a GW event is observed and limit the total time to merger to be less than
$13.7\,\rm{Gyr}$. Using the total lookback time, we can compute the redshift of
the binary at ZAMS formation following the Planck 2018 cosmological model
\citep{Planck2018}. The main benefit of post-processing the redshift in this way
is that we do not rely on a particular assumption of star formation rate (SFR)
distribution when we solve for the posterior distribution in redshift. This
means we can put the prior on formation redshift (or in general,
metallicity-redshfit distribution) in by reweighting the posterior distribution
according to a particular SFR model, and therefore do not have to rerun the
inference every time we change the SFR model. This is especially helpful since
SFR models are still highly uncertain but strongly impact the local merger rates
of BBHs \citep{Broekgaarden2021}. For hyperparameters, we choose parameters that
affect binary evolution in \texttt{COSMIC} most significantly for massive stars,
including the common envelope efficiency $\alpha$, the fraction of mass accreted
onto the accretor during stable mass transfer $f_{\rm acc}$, the critical mass
ratio which determines whether Roche-overflow mass transfer proceeds stably or
produces a common envelope when the donor star is on the Hertzsprung gap $q_{\rm
crit,3}$, and the root-mean-square of the Maxwellian distribution employed for
CO natal kicks, $\sigma$.

We show a portion of the joint posterior of the progenitor parameters and
hyperparameters of GW150914 in figure \ref{fig:GW150914_posterior}. Each panel
is colored based on the comparison between progenitor parameters (blue),
hyperparameters (orange), or a mix of the two (green). We find that the choice
for $\sigma$ does not affect the formation of GW150914-like BBHs. This is not
unexpected since we use the \citet{Fryer2012} delayed model which reduces the
natal kick strength at BH formation based on the amount of fallback onto the
proto-CO. In the case of GW150914-like BBHs, the BH progenitors are massive
enough that the fallback reduces the natal kick to zero. Because of this we do
not include $\sigma$ in Figure~\ref{fig:GW150914_posterior}. Similarly, we do
not include the ZAMS orbital period and eccentricity, which are correlated with
one another, but not strongly with the other progenitor parameters or
hyperparameters. We find strong correlations between the ZAMS masses, as well as
strong correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm{acc}}$, and
$f_{\rm{acc}}$ and $\alpha$. We show scatter plots of each of these combinations
in the upper-right inset of the figure.

The correlations between the ZAMS masses are similar to those found by A21 with
the majority of the population preferring primary and secondary masses between
$60$--$90\,M_{\odot}$. However, we also find some ZAMS masses which extend up to
$150\,M_{\odot}$ which is the limit imposed by our assumptions. We explore the
formation scenarios which lead to successful GW150914-like mergers and those
which fail to produce GW150914-like mergers below in
Figure~\ref{fig:compare_fixed_variable}.

The correlations between $M_{2,\rm{ZAMS}}$ and $f_{\rm acc}$ illustrate the
variety of ways which GW150914-like mergers can be produced. For binaries with
$M_{2,\rm{ZAMS}} < 70\,M_{\odot}$, we find that accretion efficiencies of
$f_{\rm acc} > 0.5$ are preferred. One reason for this is based on the
requirement that the total mass in the binary must remain above the total mass
of GW150914 and strongly non-conservative mass transfer (i.e. $f_{\rm acc} <
0.5$) reduces the total mass of the system. A compounding factor is that as mass
leaves the binary due to non-conservative mass transfer, the evolution of the
binary's orbit is less dramatic and leads to wider binaries on average and thus
fewer mergers in a Hubble time. For binaries with $M_{2,\rm{ZAMS}} \sim
80\,M_{\odot}$, $f_{\rm acc}$ is less constrained. This is because of a
preference for these secondaries to also have primary masses near $
80\,M_{\odot}$ and thus enter a double-core common envelope evolution in which
both stars' envelopes are ejected, leaving behind two stripped helium cores in a
tight orbit. In this case, $f_{\rm acc}$ does not affect the binary evolution
and is thus unconstrained. This effect can also be seen in the ZAMS mass and
$q_{\rm crit, 3}$ panels for masses near $80\,M_{\odot}$ because double-core
common-envelope evolution is triggered in \texttt{COSMIC} when the radii of the
two stars touch due to the rapid expansion of the primary star upon helium
ignition. In this case, the choice for Roche-overflow mass transfer to proceed
stably (as prescribed by $f_{\rm{acc}}$) or unstably (as prescribed by $\alpha$)
using $q_{\rm crit, 3}$ is totally irrelevant, and thus unconstrained. Finally,
for binaries with $M_{2,\rm{ZAMS}} > 80$, we find that $f_{\rm acc}$ is
correlated to decrease with increasing mass. This is due to limits on the total
mass which must be ejected from the binary to produce BH masses that match
GW150914, the reverse situation to $M_{2,\rm{ZAMS}} < 70 \, M_\odot$. Binaries
with both component masses near $150\,M_{\odot}$ must either go through a common
envelope evolution (in which case $f_{\rm acc}$ is unconstrained), or very
non-conservative mass transfer (where $f_{\rm acc}$ is low), to produce BBHs
with the proper mass.

Finally, we find that $\alpha$ and $f_{\rm acc}$ are largely uncorrelated,
though there exist independent trends in each hyper parameter. This is not
totally unexpected since the physical processes which are described by each
hyperparameter, i.e. stable Roche-lobe overflow and common envelope evolution
are two independent channels. Generally, we find that GW150914-like BBHs tend to
prefer larger accretion efficiencies and have common envelope ejection
efficiencies peaking near $\alpha=1$. One shortcoming of our method
implementation is the application of a single hyper parameter for
$f_{\mathrm{acc}}$ and $\alpha$ for each binary, while both the primary and the
secondary star could in principle be defined by their own hyperparameters that
prescribe the outcomes for when each star fills it's Roche lobe. We reserve full
treatment of this for future work but note that this improvement could reveal
correlations between hyperparameters that are not present in our current
analysis.

\begin{figure*}[h]
    \includegraphics[width=\textwidth]{figures/GW150914_corner_zoomed_lowres.png}
    \caption{The posterior for GW150914 in both the progenitor parameter space
    and hyperparameter space. $M_1$ and $M_2$ are the progenitors' masses.
    $\log{Z}$ is log metallicity at ZAMS. $z$ is the redshift at ZAMS. $\alpha$
    is the common envelope efficiency. $f_{\rm acc}$ is the fraction of mass
    accreted during stable mass transfer. $q_{\rm crit, 3}$ is the critical mass
    ratio on the Hertzsprung Gap. The contours correspond to the $68\%$ and
    $95\%$ confidence intervals. Note that the redshift is not fitted during the
    root finding process or the MCMC process. Once we find the evolutionary
    parameters, we add the delay time to the lookback time of the observed
    posterior sample, then from the total lookback time we can compute the
    redshift at ZAMS. We highlight three panels in the corner plots to show the
    fine structure of the set of posterior samples in the evolutionary parameter
    space. We also color the posterior in a particular panel according to the
    type of parameters involved in the corner plot. Blue denotes panels that
    include only progenitor parameters, green denotes panels that include a mix
    of progenitor parameters and hyperparameters, and orange denotes panels that
    include only hyperparameters. }
    \label{fig:GW150914_posterior}
    \script{figure1_cornerPlot.py}
\end{figure*}

Once we obtain a set of binaries which successfully map the ZAMS parameters and
hyperparameters to BBH merger masses, we re-evolve the set of ZAMS parameters
with the same physical assumptions as A21, but vary the common envelope
efficiency to explore how keeping a fixed model which only varies one
hyperparameter contrasts to our results. Figure~\ref{fig:compare_fixed_variable}
shows the distribution of ZAMS masses for three models (first through third
columns), each with a different $\alpha$ but the same hyperparameters as A21, as
well as the results of our sampling which allows our hyperparameters to vary
(fourth column). We find that holding the accretion efficiency, $f_{\rm{acc}}$,
and common envelope ejection efficiency, $\alpha$, to fixed values greatly
reduces the ZAMS parameter space that produces GW150914-like mergers. In
contrast, by allowing the hyperparameters to fully span the model uncertainty,
we find that there are distinct ZAMS parameters which produce GW150914-like
mergers.

Because we can explore the full evolutionary parameter space, we can also
determine that there are ZAMS parameters which fail to produce GW150914-like
mergers \emph{regardless of our hyperparameter choice.} For primary ZAMS masses
between $\sim95-135\,M_{\odot}$ and secondary ZAMS masses between
$\sim50-95\,M_{\odot}$, we find that there is no combination of accretion
efficiency and common envelope ejection efficiency which produces merging BBHs
that have masses consistent with GW150914's posteriors. In this region, if BBHs
with masses consistent with GW150914's masses are produced through a combination
of $\alpha$ and $f_{\rm{acc}}$, they do not merge in a Hubble time. Conversely,
in the region of the fourth column of Figure~\ref{fig:compare_fixed_variable}
which does produce GW150914-like mergers, the combination of $\alpha$,
$f_{\rm{acc}}$, and the ZAMS masses and orbital separations balance to produce
BHs with the correct masses that merge within a Hubble time.

\begin{figure*}
    \includegraphics[width=0.98\textwidth]{figures/compare_fixed_variable.png}
    \caption{Comparison of ZAMS masses for binaries which produce GW150914-like
     mergers for three variations of $\alpha$ with a fixed set of parameter
     assumptions matching those of A21 (first three columns) and for binaries
     which produce GW150914-like mergers when $\alpha$, $f_{\mathrm{acc}}$, and
     $q_{crit,3}$ are allowed to vary, (fourth column).}
    \label{fig:compare_fixed_variable}
    \script{figure2_varyAlphaFacc.py}
\end{figure*}

It is interesting to re-project the posterior over progenitor parameters
($\bm{\theta}'$) and hyperparameters ($\bm{\lambda}$) using fresh random
variables ($\bm{X}$) to the observable space of GW parameters ($\bm{\theta}$).
Agreement between the observed parameters and the re-projected parameters
indicates good exploration of the random variable space and lack of sensitivity
to the details of these random parameters.  Figure
\ref{fig:GW150914_reprojection} shows good agreement between the re-projected
root-finding outputs as well as the re-projected MCMC draws for our analysis of
GW150914.  Unlike \citet{Andrews2021}, we find there are no regions of
observable space inaccessible to our evolutionary models; this difference arises
because we allow the evolutionary hyperparameters $\bm{\lambda}$ to vary.

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GW150914_reprojection.pdf}
\caption{Reprojecting the posterior in evolutionary parameter space of GW150914
to observable space. The contours correspond to the $68\%$ and $95\%$ confidence
intervals. The blue contour is the reprojected posterior after the root-finding
procedure. The orange contour is the reprojected posterior after the MCMC
procedure. The green contour is the posterior plotted using the original LVK
posterior samples. }
\label{fig:GW150914_reprojection}
\script{figure3_sanityCheck.py}
\end{figure}

% Paragraph discussing the potential application on the population side.
To illustrate the potential benefit of using our method on the population level,
we perform the same analysis for all events in GWTC-3. Plotted in figure
\ref{fig:GWTC-3_f_acc_mass} is the posterior density in the $m_{1,{\rm
GW}}-f_{\rm acc}$ space for most of the events in GWTC-3. Each contour
represents the $68\%$ credible interval of the posterior density for that
particular event. There is a suggestive trend showing that $f_{\rm acc}$ could
increase as the mass of the progenitor increases. Such a trend implies the
stable mass transfer phase of a less massive binary would be preferentially
non-conservative, with more conservative mass transfer in more massive binary
systems.

Note that some events did not pass the KL divergence test we proposed in section
\ref{sec:method}; we do not include these systems in figure
\ref{fig:GWTC-3_f_acc_mass}. Binaries that form low-mass compact objects have
lower amounts of fallback and tend to have correspondingly larger variance in
their random natal kicks. Larger kicks can unbind the progenitor binary, or lead
to wide binaries that do not merge within a Hubble time.  In these cases, the
extra variance during the reprojection can produce a posterior that may not
agree with the original posterior, hence yielding a higher KL divergence.

For high-mass binaries, \texttt{COSMIC} struggles to produce events above the
pair instability supernova mass cutoff $m_{1,{\rm GW}}$, so the posterior in the
evolutionary parameter space only corresponds to part of the posterior in the
observable space below this cutoff, and therefore events beyond the PISN mass
gap also have higher KL divergence.

Events with more extreme mass ratios are hard to produce with \texttt{COSMIC},
and therefore less likely to be accurately recovered by our method
\citep[e.g.][]{Zevin2020}. This could be due to our method assuming a single
value for $f_{\rm acc}$ and $\alpha$ as discussed in Section~\ref{sec:result}.

Merger observations for which \texttt{COSMIC} struggles are likely to be very
informative about formation channels and binary evolution physics, and are
therefore likely worthy of close study.  We anticipate that future work will
follow up these events in detail.

Figure \ref{fig:GWTC-3_f_acc_mass} shows that our method could in principle
reveal the correlation between progenitor parameters and hyperparameters on a
population level. Here we impose a cut to eliminate events with KL divergence
larger than 0.1. This is a heuristic choice to exclude events that are obviously
not compatible with \texttt{COSMIC}. A careful treatment of all events in the
catalog and discussion related to the detailed physical implications of figure
\ref{fig:GWTC-3_f_acc_mass} is beyond the scope of this paper.  We defer a
detailed study of the physics related to the population of GW events to future
work.

\begin{figure}
\includegraphics[width=0.49\textwidth]{figures/GWTC3_f_acc_mass.pdf}
\caption{ The posterior density in the $m_{1,{\rm GW}}-f_{\rm acc}$ space for
    most of the events in GWTC-3. Note that $m_{1,{\rm GW}}$ here refers to the
    reprojected posterior instead of the posterior released by LIGO. Each
    contour is the $68\%$ credible interval of the posterior density for a
    particular event. At $m_{1,{\rm GW}} \sim 45 M_{\odot}$, the pair
    instability supernova mechanism prevents \texttt{COSMIC} from producing
    events that are more massive than this cutoff. Therefore, events with the
    majority of posterior support above this cutoff are not compatible with
    \texttt{COSMIC}, hence have a large KL divergence and are excluded from this
    figure. On the low mass end, neutron star binaries or neutron star-black
    hole binaries are subjected to randomness induced by the natal kick, also
    resulting in a larger and fluctuating KL divergences, and therefore are also
    excluded from the analysis. }
\label{fig:GWTC-3_f_acc_mass}
\script{figure4_population.py}
\end{figure}

\section{Discussion}
\label{sec:discussion}

We present a new pathway to understand binary evolution with GW events in this
paper. Instead of forward modelling an assumed distribution of initial binaries
to the observed population, we find the corresponding evolutionary parameters
event-by-event. In this first work, we showcase the power of the proposed method
with an application to GW150914. We show the joint posterior of the event's
progenitor parameters and hyperparameters.  Our work is both a more efficient
way to study GW event progenitors, and also allows the possibility of
constraining the astrophysics related to binary evolution, especially by
capturing the correlation between hyperparameters in different systems.

Our results suggest that the accretion efficiency during stable mass transfer
may depend on the primary black hole mass.  In general, our method returns the
joint posterior distribution of progenitor parameters and hyperparameters for
each event, which enables a data-driven way to study the distribution of
hyperparameters.  That is, once we have a catalog of events, each has their own
posterior in the hyperparameter space, and we can employ well-known techniques
such as hierarchical Bayesian analysis to fit a population model to the
distribution of hyperparameters; this avoids making overly-specific assumptions
like fixing the hyperparameters for all types of event.

% Rate analysis is now just counting, like what's done in phenomenological
% model, instead of forward modelling.
Working event-by-event as we do permits post-processing application of
arbitrarily complicated models of star formation and metalliticy evolution by
re-weighting the samples we obtain \citep[e.g.][]{VanSon2022}. Once we have
pulled the GW event posterior back from the observable space to the evolutionary
parameter space, we can apply the same hierarchical inference methods currently
used to infer the compact binary population to our progenitor population.  In
particular, we can use our progenitor population to infer the formation rate of
compact binary population progenitor systems over cosmic time without needing to
make artifical and simplistic assumptions about the delay time distribution or
the metallicity-specific star formation rate
\citep{Vitale2019,Ng2021,VanSon2022}.  Comparing the formation rate of
progenitors to the star formation rate provides an avenue to check our
understanding of binary evolution and the relative rate of other formation
channels.

% Single formation channel and fundamental limitation on that.
While our method allows data-driven exploration of the hyperparameters space for
the first time, there are a number of improvements that can be implemented in
future studies. In this study we use only \texttt{COSMIC} as our evolutionary
function, which by design cannot explain all the events in GWTC-3. For example,
event with either of the component mass larger than the lower edge of PISN gap
such as GW190521 cannot be explained by \texttt{COSMIC}. Alternative channels
such as dynamical formation might be needed to explain some subset of the events
in GWTC-3 \citep{Zevin2021}. As the sensitivity of GW detectors increases, we
expect to see more and more events that are unusual in some way. Therefore,
having a self-consistence population synthesis code that contains multiple
formation channels is essential to accommodate the growing catalog of GW events.
In this paper, our main focus is to illustrate the concept of "back-propagating"
GW events posterior samples, highlighting the capability of our method and
motivating the community to build the next generation population synthesis code
that can work with our method. To avoid cluttering of focus, we discuss the
physical implication of the result presented in this work under our specific
assumption (i.e. using \texttt{COSMIC} as our evolutionary function) in a
companion paper.

% Randomness paragraph
Due to the implicit definition of random variables in \texttt{COSMIC}, our
evolutionary function is stochastic. This introduces significant inefficiency in
our root-finding and sampling algorithm. The main stochasticity in
\texttt{COSMIC} comes from natal kick, which significantly affect the
evolutioary pathway of low mass events such as BNS mergers. The effect of the
natal kick is suppressed for heavier mass events due to fallback. This means
events with lighter masses are subject to stochasticity of the function, where
the sampling process for heavier events behaves as if the evolutionary function
we use is deterministic.
% Indeed, we see this behavior in a couple of case studies. For heavy event such
% as GW150914, the reprojected posterior agrees with the data posterior. For
% light event such as GW170817, although during the root-finding process the
% convergence criteria is met, the reprojected posterior does not agree with the
% data posterior.
Due to computational limitation, we only try 1000 different initial guesses per
posterior sample in the root-finding process. This means any posterior sample
that has a probability of merging rarer than 1 in 1000 could be missed.
Obviously the problems which comes with the randomness can be alleviated by
performing more tries per posterior sample, but this is not scalable in
practice. On top of limitation in efficiency, some formation channels require
explicit control of random variables by construction. For example, in a
dynamical formation scenario such as binaries that form in a globular cluster,
each binary has some probability of undergoing a multi-body encounter with
another member in the cluster. These encounter probability distributions are
either studied with direct N-body simulations or semi-analytical methods. In
both cases, each member of the cluster is no longer completely independent of
the other members, but coupled through the encounter probability distribution.
By studying the encounter probability distribution, we can infer the properties
of the environment which the binary lives in. This can only be done if we have
explicit control over the random variables that characterize the encounter
probability distribution.

% Need for selection bias to recover intrinsic distribution

% Gradient decending in the evolutionary parameter space is much more efficent
% than rejection sampling. Potential pitfall of interpreting this result. Need
% for full autodiff.

% By leveraging gradient-based optimization strategy, obtaining posterior
% samples in the evolutionary parameter space is much more efficient than
% rejection sampling. In each optimization step during the root-finding process,
% the current state of the guess and its gradient provide useful information
% about where the root might be. In contrast to sampling algorithm, which
% discard the sample whenever it is not accepted. This greatly benefit the
% efficiency of the root-finding process. On the other hand, this method is not
% a sampling algorithm, but a coordinate transform of the posterior samples. One
% should pay extra caution in interpreting the posterior samples in the
% evolutionary parameter space. Since posterior samples in the observable space
% can in principle have multiple roots in the evolutionary parameter space, the
% relatively weighting of these roots becomes ambiguous when the number of roots
% is large. For example, should we consider roots that are close to each other a
% single root or multiple roots? If we have two clusters of roots, the relative
% number of roots within each cluster would determine the relative weight of
% each mode. Note that since they both map to the same observable posterior
% sample within the acceptance threshold, all weighting assigned are equally
% valid from a "matching data" standpoint. This leave the interpretation of the
% posterior in the evolutionary parameter space ambiguous. In our case, we found
% empirically each posterior sample may at most only correspond to a handful of
% roots (most of the posterior samples have a unique root.), therefore we choose
% to assign equal weight to each root.

% The potential reason for the good agreement between MCMC and the root finding
% is due to the behavior of posterior in the evolutionary parameter space are
% larger determined by the posterior distribution in the observable space
% instead of multiple possible roots per posterior sample. We have also compared
% two MCMC results, one is initialized using the roots and another is
% initialized using samples drawn from a uniform prior. We found empirically the
% MCMC result using the roots as initialization converges at least an order of
% magnitude faster than the uniform prior initialization. This makes sense since
% the roots basically eliminate the long burn-in needed to go from random
% location in the evolutionary parameters to the most probable region. We leave
% a detail comparison between this method and direct sampling to future studies.

Another technical note is that we use finite differencing to estimate the
gradient of the objective function, which could be a significant source of error
near transition points in the evolutionary parameter space. Also, finite
differencing is increasingly inefficient as we increase the dimensionality of
the problem. To improve the accuracy and efficiency in estimating the gradient
of the objective function, automatic differentiation is a promising feature that
modelers should consider incorporating in their population synthesis codes in
the future.

To summarize, we propose a novel method to recover the posterior samples in the
evolutionary parameter space for each GW event. We point out hyperparameters in
the usual population synthesis simulation context are not actually parameters
related to the population, but parameters about the evolutionary function. This
means the binary evolution functions can be constrained on an individual event
basis. We "back propagate" the posterior in the observable space to the
evolutionary parameter space, thus allowing us to study hyperparameters and
theirs correlations with progenitor parameters in a data-driven manner. Our
method makes less assumptions than the traditional forward modelling approach,
which often fix the hyperparameters across the entire population. Since we are
not limited to the fixed hyperparameters assumptions, we can explore the
behavior of the hyperparameters across the population much more efficiently.
While our work lays down a data analysis pathway to understand the population of
GW events, no physics can be learned without a comprehensive physical model. We
hope this letter will motivate the community to build the next generation
population synthesis codes that have the following properties: first, they
should have explicit control over the random variables so marginalizing over
random variables can be done more precisely; second, they should be as
automatically differentiable as much as possible so exploring the evolutionary
parameter space is efficient. Combining this work and the next generation of
population synthesis codes, we can explore the full parameter space of binary
evolution models with the next-generation GW detectors network in the near
future.

\section{Acknowledgement}
The authors are grateful for discussions with the CCA Gravitational Wave
Astronomy Group. The Flatiron Institute is supported by the Simons Foundation.

\software{ \texttt{corner}\ \citep{corner}; \texttt{COSMIC}\
    \citep{Breivik2020}; \texttt{julia}\ \citep{Julia-2017} \texttt{matplotlib}\
    \citep{matplotlib}; \texttt{numpy}\ \citep{numpy}; \texttt{pandas}\
    \citep{McKinneyPandas, reback2020pandas}; \texttt{scipy}\ \citep{scipy}
    \texttt{seaborn}\ \citep{Waskom2021} \texttt{showyourwork}\
    \citep{Luger2021} }

\bibliography{bib}

\end{document}